A Bidirected String Graph Model for Genome Assembly

### 2012-12-17 presentation/paper.tex
circle colors
paper: add missing "be"

\title{A Bidirected String Graph Model for Genome Assembly}
\author{Eric Biggers}

@@

A \KeyTerm{walk} in a bidirected graph $G$ is a continuous sequence of edges in
$G$ such that if any vertex $v$ is entered through a head inwards, it is exited
through a head outwards (unless it is the end of the path), and vice versa.  An
equivalent formulation is that if we consider a possible walk in a bidirected
graph, such as the possible walk $1 \to 2 \to 3 \to 4$ in the graph shown in
\Figure{fig:bidiwalk_example}, in order to check whether the walk is valid or
not, we must consider every vertex in the possible walk except for the first and
last vertices.  For each such vertex, we need to verify that each time that
vertex is passed through as part of a walk, the arrow heads adjacent to that
vertex that are used in the relevant part of the walk are pointed in the same
direction {\em relative to the walk}. If there are no such pairs of heads that
point towards each other or point away from each other, the walk is valid.  The
red circles in \Figure{fig:bidiwalk_example} show the two pairs of heads that
need to be checked for the possible walk $1 \to 2 \to 3 \to 4$.

@@

\begin{figure}
	\begin{center} {\small
		\begin{tikzpicture}[>=triangle 45]
			\tikzstyle{every node} = [circle,fill=\nodecolor,draw=black];
			\node (1) at (0.00, 0) {1};
			\node (2) at (2.00, 0) {2};
			\node (3) at (4.00, 0) {3};
			\node (4) at (6.00, 0) {4};
			\draw[<->,very thick] (1) edge (2);
			\draw[>-<,very thick] (2) edge (3);
			\draw[<-<,very thick] (3) edge (4);
			\draw[color=red,thick] (2.00, 0) circle (0.8);
			\draw[color=red,thick] (4.00, 0) circle (0.8);
		\end{tikzpicture}
	} \end{center}
	\caption{$1 \to 2 \to 3 \to 4$ and $4 \to 3 \to 2 \to 1$ are both valid
	walks in the bidirected graph above.  (Check circled pairs of heads to
	verify this.)}
	\label{fig:bidiwalk_example}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=1.0\textwidth]{E_coli.pdf}
		\caption{Transitively-reduced, collapsed bidirected string graph for
		1000 bp perfect reads uniformly sampled every 100 bp from random strands
		of the non-circular version of the 4,639,675 bp genome of {\it E. coli}
		(a common bacterium) using a minimum overlap length of 250 bp.  This
		graph was produced by the code written for this project, but the actual
		drawing of the picture was done by {\it
		GraphViz}\cite{GraphVizPublication}.  The 46,387 reads
		that were provided as input to the assembler were generated by a Perl
		script.  The bottleneck of the algorithm up to this point is
		\ProgramName{compute-overlaps}, which took about 10 minutes and used a
		maximum of about 1 GiB of memory to find all 336,227 overlaps from these
		reads, seeding on 128-mers.  The running time and memory usage of
		\ProgramName{compute-overlaps} would need to improved further to handle
		larger genomes.}
		\label{fig:E_coli}
	\end{center}
\end{figure}

>>>>

\title{A Bidirected String Graph Model for Genome Assembly}
\author{Eric Biggers}
\date{December 12, 2012}

@@

A \KeyTerm{walk} in a bidirected graph $G$ is a continuous sequence of edges in
$G$ such that if any vertex $v$ is entered through a head inwards, it is exited
through a head outwards (unless it is the end of the path), and vice versa.  An
equivalent formulation is that if we consider a possible walk in a bidirected
graph, such as the possible walk $1 \to 2 \to 3 \to 4$ in the graph shown in
\Figure{fig:bidiwalk_example}, in order to check whether the walk is valid or
not, we must consider every vertex in the possible walk except for the first and
last vertices.  For each such vertex, we need to verify that each time that
vertex is passed through as part of a walk, the arrow heads adjacent to that
vertex that are used in the relevant part of the walk are pointed in the same
direction {\em relative to the walk}. If there are no such pairs of heads that
point towards each other or point away from each other, the walk is valid.  The
circles in \Figure{fig:bidiwalk_example} show the two pairs of heads that need
to be checked for the possible walk $1 \to 2 \to 3 \to 4$.

@@

\begin{figure}
	\begin{center} {\small
		\begin{tikzpicture}[>=triangle 45]
			\tikzstyle{every node} = [circle,fill=\nodecolor,draw=black];
			\node (1) at (0.00, 0) {1};
			\node (2) at (2.00, 0) {2};
			\node (3) at (4.00, 0) {3};
			\node (4) at (6.00, 0) {4};
			\draw[<->,very thick] (1) edge (2);
			\draw[>-<,very thick] (2) edge (3);
			\draw[<-<,very thick] (3) edge (4);
			\draw[color=black!70] (2.00, 0) circle (0.8);
			\draw[color=black!70] (4.00, 0) circle (0.8);
		\end{tikzpicture}
	} \end{center}
	\caption{$1 \to 2 \to 3 \to 4$ and $4 \to 3 \to 2 \to 1$ are both valid
	walks in the bidirected graph above.  (Check circled pairs of heads to
	verify this.)}
	\label{fig:bidiwalk_example}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=1.0\textwidth]{E_coli.pdf}
		\caption{Transitively-reduced, collapsed bidirected string graph for
		1000 bp perfect reads uniformly sampled every 100 bp from random strands
		of the non-circular version of the 4,639,675 bp genome of {\it E. coli}
		(a common bacterium) using a minimum overlap length of 250 bp.  This
		graph was produced by the code written for this project, but the actual
		drawing of the picture was done by {\it
		GraphViz}\cite{GraphVizPublication}.  The 46,387 reads
		that were provided as input to the assembler were generated by a Perl
		script.  The bottleneck of the algorithm up to this point is
		\ProgramName{compute-overlaps}, which took about 10 minutes and used a
		maximum of about 1 GiB of memory to find all 336,227 overlaps from these
		reads, seeding on 128-mers.  The running time and memory usage of
		\ProgramName{compute-overlaps} would need to be improved further to handle
		larger genomes.}
		\label{fig:E_coli}
	\end{center}
\end{figure}
